---
title: "Yukon Canadian-origin Chinook Salmon Inseason Bayesian Daily Projection"
author: "Aaron Lambert"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    theme:
      bootswatch: cerulean
---

# July 19, 2023
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, results = FALSE, warning = FALSE)
library(kableExtra)


```



```{r, echo=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)
require(rstan)
require(rstanarm)
require(bayesplot)
require(tidyverse)
require(ggthemes)
require(viridis)
require(shinystan)
require(lubridate)
require(ggpubr)
require(gridExtra)
require(tidybayes)
library(wesanderson)
require(grid)
library(knitr)
# set to working directory
wd <- "C:/Users/aaron/Desktop/Yukon Kings/Inseason Forcast Model"

setwd(wd)

# Objects used to save/load data, outputs, or stan/R scripts
dir.output <- file.path(wd,"output")
dir.figs <- file.path(wd,"figs")
dir.stan <- file.path(wd,"Stan")
dir.data <- file.path(wd,"Data")
dir.R <- file.path(wd,"R")

# Read in the file as xlsx !!!CHANGE FILE HERE!!!!!!!!!
PSS <- read_xlsx(file.path(dir.data,"ADFG PSS Daily/Yukon Escapement Daily 19July23.xlsx"),skip = 3)

ESS <- read_xlsx(file.path(dir.data,"ADFG Eagle Daily/Yukon Escapement Daily Eagle 19July23.xlsx"),skip = 3)

# Add a day column with the appropriate day ranges
PSS$Day <- c(148:252)
ESS$Day <- c(178:243)

# PSS <- as.data.frame(PSS)

# Convert to long format 
test2 <- PSS[,2:30] %>% 
  pivot_longer(cols = -c(Day), names_to = "Year", values_to = "count") %>% 
  as.data.frame()

testEagle <- ESS[,2:21] %>% 
  pivot_longer(cols = -c(Day), names_to = "Year", values_to = "count") %>% 
  as.data.frame()

# replace NA's with 0
test2[is.na(test2)]<-0

testEagle[is.na(testEagle)]<-0

# Change Year to numeric
test2$Year <- as.numeric(test2$Year)

testEagle$Year <- as.numeric(testEagle$Year)

# Arrange data set by year
PSS_hist<- test2 %>% arrange((Year))

Eagle_hist<- testEagle %>% arrange((Year))

# make sure it worked
head(PSS_hist)
dim(PSS_hist)

# Clean up the environment
rm(test2,PSS, ESS, testEagle)



# Parralize for optimum model run time
rstan_options(auto_write = TRUE)
#
mc.cores = parallel::detectCores()
# mc.cores <-1

# # Call in functions from model_run_funciton.r
source(file = file.path(dir.R,"model_run_function LOO3.r"))

######### Import Data ##############################################################
# Historical Canadian EOS reconstructed run
# This is the reconstructed data from Curry for old reconstructed modeling procedure
CAN_hist <- readRDS(file.path(dir.data,"Canadian Passage RR 21Mar23.RDS"))

# PSS historical for Jun 1st = 152, to Aug 31 = 243
# Data obtained from ADFG website
# https://www.adfg.alaska.gov/index.cfm?adfg=commercialbyareayukon.salmon_escapement
# NOTE!!!! See use "Dataframe preprocess.R" to get  most recent version.
#  If not using Dataframe preprocess.R, uncomment out the following to load relevant PSS historical data
# PSS_hist <- readRDS(file = file.path(dir.data,"PSS passage 22Jun22.RDS"))

# Read in historical avg of GSI by strata (Naive estimator)
# GSI_mean <- readRDS(file = file.path(dir.data,"Mean GSI by strata 2005-2020.RDS"))

# Read in historic preseason forecasts (2013 - current)
# pf_hist <- readRDS(file.path(dir.data,"preseason forcast.RDS"))
pf_hist <- readRDS(file.path(dir.data,"pf_ver3.1_12June23_eagle+harvest.RDS"))
# Read in genetic stock identification (2005-2019) 
# (adjusted to capture early and late runs)
# GSI_by_year <- readRDS(file = file.path(dir.data,"GSI by year"))
GSI_by_year <- readRDS(file = file.path(dir.data,"GSI by year unadj 21Mar23.RDS"))

# Shape parameters for fitting curves
logistic.all <- read.csv(file = file.path(dir.output,"logistic curve parameters All Chinook 1995_2022.csv"))

# Normal curve fitting parameters
normal.all <- read.csv(file = file.path(dir.output,"normal curve parameters All Chinook 1995_2022.csv"))

# Read in PSS observation error estimates
PSS_sd <- readRDS(file = file.path(dir.data,"PSS SD 1995_2021.RDS"))

# SST May data
norton.sst <- readRDS(file = file.path(dir.data,"norton.sst 6Apr23.RDS"))

# Control Section ##################################################################
model.version <- "4.0.2"

# Range of years 1995 to 2022 (1996 is not available)
myYear <- 2023
# 
# # Range of days 152 (June 1) - 243 (August 31)
# # myDay <- 171
# 
# # Set my day using the myDay_func function
# #  Enter month and day as numbers
myDay <- myDay_func(Month = 7, Day = 19)

# Days for plotting date
testDays <- seq(from = 153, to = 213, by = 5)

# MCMC Parameters
n.chains <- 4
n.iter <- 7000;#5e4
n.thin <- 2


# Run the model for a single year and day of interest.
#  Ajust the inputs for the model in the control section above
model.output <-InSeasonProjection(model.version = model.version,
                                  myYear = myYear,
                                  myDay = myDay,
                                  n.chains = n.chains,
                                  n.iter = n.iter,
                                  n.thin = n.thin,
                                  CAN_hist = CAN_hist,
                                  PSS_hist = PSS_hist,
                                  # GSI_mean = GSI_mean,
                                  pf_hist = pf_hist,
                                  GSI_by_year = GSI_by_year,
                                  savefit = TRUE, 
                                  Eagle_hist = Eagle_hist,
                                  normal = TRUE,
                                  logistic = FALSE,
                                  prior.df.log = logistic.all,
                                  prior.df.norm = normal.all,
                                  multiplier = 1,
                                  startDayPSS = 148,
                                  startYearPSS = 2005,
                                  PSS_sd = PSS_sd,
                                  norton.sst = norton.sst)

run.size.ecdf <- ecdf(model.output$pars$RunSize)

prob.escapement <- run.size.ecdf(42500)

```

Today's forecast using the Bayesian inseason projection for Canadian-origin Chinook salmon returning to the Yukon River is found below. Given the preseason forecast and the observed Chinook salmon passage at PSS and Eagle Sonar up to July 19, 2023, there is a `r (1-prob.escapement)*100` % chance of meeting the lower range of the interim management escapement goal of 42,500-55,000 Canadian-origin Chinook salmon.

For this forecast, model version 4.0.2 is used (Normal curve fit to daily PSS passage and Eagle proportion est).


# Projected runsize quantiles
```{r pressure, echo=FALSE, results='markup', fig.align='left'}
quant <-round(quantile(x = model.output$pars$RunSize, probs = c(.025,.1,.5,.9,.975)))

tble <- data.frame("Quantile"  = names(quant), "Run size" = quant)

rownames(tble) <- NULL

kable(tble,
      # align = "crr", 
      col.names = c("Quantile", "Run Size")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                position = "center", font_size = 18,
                latex_options = "hold_position")

```


\newpage

# Cumulative PSS Chinook Passage (2005-2023)
Observed cumulative PSS passage from years 2005-2023. Note that the current year is plotted in black.
```{r,fig.align='center'}
cumcount <- PSS_hist[PSS_hist$Year>=2005 &
                       PSS_hist$Day<=225,] %>% group_by(Year) %>% mutate(CumCount = cumsum(count)) %>% as.data.frame

ggplot(cumcount[cumcount$Year < 2023,], aes(x = Day, y = CumCount/1000, color = factor(Year)))+
  # geom_col(fill = "red")+
  geom_line()+
  geom_point(data = cumcount[cumcount$Year==myYear &
                               cumcount$Day<= myDay,], aes(x = Day, y = CumCount/1000), color = "black")+
  xlab("Day of Year")+
  ylab("Thousands of Chinook Salmon")+
  scale_x_continuous(breaks = c(testDays), labels = c(
    "June 2",
    "June 7",
    "June 12",
    "June 17",
    "June 22",
    "June 27",
    "July 2",
    "July 7",
    "July 12",
    "July 17",
    "July 22",
    "July 27",
    "Aug 1"))+
  labs(color = "")+
  # theme_tidybayes()+
  theme(legend.position = "top",
        panel.grid.major.y  =element_line(linetype = 2, size = 1, 
                                          color = "grey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA),
        plot.background = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 12),
        axis.text = element_text(angle = 90))
```
\newpage

# Cumulative Eagle Sonar Chinook Passage (2005-2023)
Observed cumulative Eagle sonar passage from years 2005-2023. Note that the current year is plotted in black.
```{r,fig.align='center'}

cumcount_eagle <- Eagle_hist[Eagle_hist$Year>=2005 &
                       Eagle_hist$Day<=225,] %>% group_by(Year) %>% mutate(CumCount = cumsum(count)) %>% as.data.frame

ggplot(cumcount_eagle[cumcount_eagle$Year < 2023,], aes(x = Day, y = CumCount/1000, color = factor(Year)))+
  # geom_col(fill = "red")+
  geom_line()+
  geom_point(data = cumcount_eagle[cumcount_eagle$Year==myYear &
                               cumcount_eagle$Day<= myDay,], aes(x = Day, y = CumCount/1000), color = "black")+
  xlab("Day of Year")+
  ylab("Thousands of Chinook Salmon")+
  scale_x_continuous(breaks = c(testDays), labels = c(
    "June 2",
    "June 7",
    "June 12",
    "June 17",
    "June 22",
    "June 27",
    "July 2",
    "July 7",
    "July 12",
    "July 17",
    "July 22",
    "July 27",
    "Aug 1"))+
  labs(color = "")+
  # coord_cartesian(xlim = c(178,200),
  #                 ylim = c(0,20))+
  theme(legend.position = "top",
        panel.grid.major.y  =element_line(linetype = 2, size = 1, 
                                          color = "grey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA),
        plot.background = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 12),
        axis.text = element_text(angle = 90))
```

\newpage



# Arrival Distribution
Curve fit to observed daily PSS Chinook salmon passage. This curve is used to project the number of Chinook salmon that will pass PSS at the end of the season to be used to project the end-of-season Canadian-origin Chinook salmon run size using a regression relationship. The average midpoint of the run from years 2005-2022 is plotted as a black dashed line.

```{r, fig.align='center'}
quant.predPSS.curve <- apply(X = model.output$pars$ps_pred_curr, 
                       MARGIN = 2, 
                       FUN = quantile, 
                       probs=c(0.025, 0.25, 0.5, 0.75, 0.975))



pred.plot.df <- data.frame(model.output$dayPSS_all,
                           c((PSS_hist$count[PSS_hist$Year == myYear & PSS_hist$Day <= myDay]),rep(x = 0,250-(myDay))),
                           t(quant.predPSS.curve))

names(pred.plot.df) <- c("Days","cumcount","low95","low50","median",
                         "up50","up95")

pred.plot.df$Obs <- ifelse(pred.plot.df$Days <= myDay, "Obs", "Unobs")

pred.plot.df$Year <- myYear

avg.MP <- mean(normal.all$mid[normal.all$year>=2005])

pred.plot.df$Avg.MP <- avg.MP

pred.plot.df$yscale <- c(0, rep(max(pred.plot.df$up95), times = length(pred.plot.df$Avg.MP)-1))
# Plot results
ggplot(pred.plot.df, aes(x = Days,  y = cumcount/1000))+
  geom_point(aes(color = Obs), size = 2) +
  # geom_col(alpha = .3)+
  geom_ribbon(aes(ymin=low95/1000, ymax=up95/1000,fill = "HDI 95"), alpha=0.3) +
  geom_ribbon(aes(ymin=low50/1000, ymax=up50/1000, fill = "HDI 50"), alpha=0.3) +
  geom_line(aes(y=median/1000, color = "Median"), show.legend = T) +
  coord_cartesian(x = c(148,215))+
  labs(x = "Day",
       y = "Chinook Salmon (1000's)")+
  geom_line(aes(x = avg.MP, y =  yscale/1000, color = "Avg Mid-point"), linetype = 2)+
  scale_x_continuous(breaks = c(testDays), labels = c(
    "June 2",
    "June 7",
    "June 12",
    "June 17",
    "June 22",
    "June 27",
    "July 2",
    "July 7",
    "July 12",
    "July 17",
    "July 22",
    "July 27",
    "Aug 1"#,
    
  ))+
  # facet_wrap(~)+
  theme(legend.position = "top",
        panel.grid.major.y  =element_line(linetype = 2, size = 1, 
                                          color = "grey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA),
        plot.background = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 14),
        axis.text = element_text(angle = 90))+
  
  scale_fill_colorblind(name = "")+
  scale_color_colorblind(name = "")+
  scale_linetype_binned()
```

\newpage

# PSS Regression Fit
Regression fit to the end-of-season cumulative PSS passage against historic observed Canadian-origin Chinook run size. Note that the current years prediction is based on summing up the observed passage up to the day of the forecast and adding the interpolated Chinook passage for the remainder of the season as predicted by fitting the normal distribution curve to observed passage.

```{r,  fig.align='center'}
# Get the quantiles
(quant_pss <-round(quantile(x = model.output$pars$curr_predPSS, probs = c(.025,.1,.5,.9,.975))))

# Get the PSS passage
tt <- sum(PSS_hist$count[PSS_hist$Year==2023])/1000

# Put everything into a dataframe for plotting
predDF <- data.frame("quant" = c(quant_pss/1000), "PSS" = rep(tt, length(quant_pss)),
                     "percent" = c("2.5%","10%","50%","90%","97.5%"))
predDF$percent<- as.factor(predDF$percent)

predDF

trial <- seq(from = 0, to = 300000, by = 1000)
length(trial)
reg.mat <- matrix(nrow = length(model.output$pars$alpha),
                  ncol = length(trial))

for(i in 1:length(trial)){
  
  reg.mat[,i] <- trial[i]*model.output$pars$beta + model.output$pars$alpha
}


quant.predPSS.beta <- apply(X = reg.mat, 
                       MARGIN = 2, 
                       FUN = quantile, 
                       probs=c(0.025, 0.25, 0.5, 0.75, 0.975))

# Get order of cumPSS for polygon below
cumPSS_all <- model.output$cumPSS_all

totalEOS <- model.output$totalEOS

ord <- order(trial)

yearPSS <- c(2005:2022)
  
# Regression plot for non-GSI-adjusted PSS passage
plot(x = cumPSS_all/1000, y = totalEOS/1000,
     type = "p",
     pch = 21,
     bg = "red",
     xlab = paste("Total EOS PSS Chinook Salmon (1000's)" ),
     ylab = ("Total EOS Canadian Chinook Salmon (1000's)"),
     # ylab = paste("1000's Canadian Chinook Salmon (1995-",myYear-1,")"),
     # main = paste("Predicted PSS Fit","1995 to", 
     # myYear-1,"\n Day 152 to",
     # myDay,"\n Model Version",model.version),
     xlim = c(0,(max(cumPSS_all/1000)+20)),
     ylim = c(min(quant.predPSS.beta/1000)-30, max(quant.predPSS.beta/1000)+30),
     # ylim = c(0,350),
     cex.axis = 1.2,
     cex.lab = 1.2)
polygon(x = c(trial[ord]/1000, rev(trial[ord]/1000)),
        y = c(quant.predPSS.beta[1,ord]/1000,rev(quant.predPSS.beta[5,ord]/1000)),
        col=rgb(1,0,0, alpha=0.2), border=FALSE)
polygon(x = c(trial[ord]/1000, rev(trial[ord]/1000)),
        y = c(quant.predPSS.beta[2,ord]/1000,rev(quant.predPSS.beta[4,ord]/1000)),
        col=rgb(1,0,0, alpha=0.2), border=FALSE)
lines(x = trial[ord]/1000, quant.predPSS.beta[3,ord]/1000, col = "red", lw = 2)
segments(x0 = mean(model.output$pars$cum_current_PSS)/1000,
         x1 = mean(model.output$pars$cum_current_PSS)/1000,
         y0 = 0,
         y1 = predDF$quant[predDF$percent=="50%"], 
         lwd = 2, 
         col = "blue")
segments(x0 = 0,
         x1 = mean(model.output$pars$cum_current_PSS)/1000,
         y0 = predDF$quant[predDF$percent=="50%"],
         y1 = predDF$quant[predDF$percent=="50%"], 
         lwd = 2, 
         col = "blue")
points(x = mean(model.output$pars$cum_current_PSS)/1000,
       y = predDF$quant[predDF$percent=="50%"],
       pch = 22,
       bg = "blue",
       cex = 2)
grid()
text(x = cumPSS_all/1000, y = (totalEOS/1000)+10, labels = c(2005:2022))
text(x = mean(model.output$pars$cum_current_PSS)/1000, y = predDF$quant[predDF$percent=="50%"]+20, "2023", col = "blue")
legend("bottomright",
       legend = c("Historic Years",
                                "This Years PSS prediction"),
       pch = c(21,22),
       col= c("red","blue")
       )

```
\newpage


# Posterior density plots 
Posterior density plots for EOS Canadian-origin Chinook salmon run size based on the preseason forecast point estimate and the empirical past performance, PSS sonar regression using predicted EOS PSS passage as the predictor variable, and Eagle Sonar passage based on the average observed proportion on the day the projection is made. Note that predictions based on Eagle Sonar are not used until Chinook salmon passage is observed at the border.


```{r, fig.align='center', fig.width=8}
# outPlots(outputList = model.output ,
#          CAN_hist = CAN_hist, GSI = FALSE, Retrospective = FALSE, eagle = TRUE)[[]]+
#   xlab("Thousands of Chinook")+
#   ylab("Relative Probability")

# Posterior EOS Can-orig run size estimate
df.run <- data.frame("par" = "Runsize",
                     "value" = model.output$pars$RunSize)

# PF (prior)
df.prior <- data.frame("par" = "Prior", 
                       "value" = model.output$pars$prior_pf)

# PSS prediction density
df.postPredPss <- data.frame("par" = "PSS Prediction", 
                             "value" = model.output$pars$post_curr_predPSS)

# Eagle prediction density
df.postPredEagle <- data.frame("par" = "Eagle Prediction", 
                               "value" = model.output$pars$post_curr_predEagle)

# Bind into one data frame with 2 columns for plotting
df.comb <- rbind(df.prior,
                 df.postPredPss,
                 df.postPredEagle,
                 df.run)
str(df.comb)
df.comb$par <- as.factor(df.comb$par)

df.comb$par <- relevel(df.comb$par,"Prior")

# Density plots comparing pf, linear prediction, and posterior estimate
# ggplot(df.comb[df.comb$par != "Eagle Prediction", ], aes(x = value/1000, fill = par))+
ggplot(df.comb, aes(x = value/1000, fill = par))+
  geom_density(alpha = .5, n = 1e5)+
  ggtitle(paste("Density plot for",myYear,", day",myDay,"\n Version", model.version))+
  theme_classic()+
  # ylim(c(0,500))
  labs(x = "1000's of Chinook Salmon", 
       fill = "Parameters", 
       y = "Probability Density")+
  # scale_fill_discrete(name = "Parameters", 
  # labels = c( "Preseason Forecast (Prior)", 
  # "PSS Prediction","Runsize"))+
  scale_x_continuous()+
  scale_fill_colorblind(name = "",
                        labels = c( "Preseason Forecast (Prior)",
                                    "Eagle Prediction",
                                    "PSS Prediction",
                                    "Runsize"))+
  # theme_tidybayes(element_text(size = 20))+
  theme(text = element_text(size = 16, family = "serif"),
        legend.text = element_text(size = 12),
        panel.grid.major =element_line(),    #strip major gridlines
        panel.grid.minor = element_blank(),    #strip minor gridlines
        legend.position = "top",
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))+
  scale_color_discrete(name = "")+
  coord_cartesian(xlim = c(0,65))

```
\newpage

```{r, fig.align='center', fig.width= 8}

ggplot(df.comb, aes(x = value/1000, y =fct_rev( par)))+
  stat_eye(aes(fill = after_stat(level)), .width = c(.8,.95,1), alpha = .5)+
  labs(x = "1000's of Chinook Salmon", 
       fill = "Parameters", 
       y = "Parameter")+
  scale_fill_colorblind(name = "Credible Interval",
  labels = c( "100%",
  "95%","80%"))+
  scale_x_continuous()+
  # scale_fill_colorblind(name = "", 
  #                       labels = c( "Preseason Forecast (Prior)",
  #                                   "PSS Prediction",
  #                                   "Runsize"))+
  theme_tidybayes()+
  theme(text = element_text(size = 20, family = "serif"),
        legend.position = "top",
        panel.grid.major =element_line(),    #strip major gridlines
        panel.grid.minor = element_blank(),    #strip minor gridlines
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))+
  scale_color_discrete(name = "")+
  guides(fill = guide_legend(override.aes = list(size = 5)))+
  coord_cartesian(xlim = c(0,150))
```